gratia: An R package for exploring generalized additive models

### 2024-06-27 paper.Rmd
change title

---
title: 'gratia: An R package for working with generalized additive models'
tags:
  - R
  - splines
  - generalized additive models
  - posterior simulation
authors:
  - name: Gavin L. Simpson
    orcid: 0000-0002-9084-8413
    affiliation: 1
affiliations:
 - name: Department of Animal and Veterinary Science, Aarhus University, Denmark
   index: 1
date: 26 June 2024
citation_author: Simpson
year: 2024
bibliography: paper.bib
output:
  rticles::joss_article:
    keep_md: true
csl: apa.csl
journal: JOSS
---

@@

Model diagnostic plots can be produced using `appraise()`, which by default produces four plots: i) a QQ plot of model residuals, with theoretical quantiles and reference bands generated following @Augustin2012-sc, ii) a plot of residuals (deviance residuals are the default) against linear predictor values, iii) a histogram of residuals, and iv) a plot of observed versus fitted values. Model diagnostic plots for the model, with simulated residuals-based reference bands on the QQ plot, are produced with
```{r m1-appraise, fig.show = "hide", dependson = "fit-chl-model", cache = FALSE, fig.height = 7, fig.width = 10, fig.pos = "t!"}
appraise(m1, method = "simulate")
```
which show significant heteroscedasticity and departure from the condtional distribution of the response given the model (Figure \ref{fig:m1-appraise}).

>>>>

---
title: 'gratia: An R package for exploring generalized additive models'
tags:
  - R
  - splines
  - generalized additive models
  - posterior simulation
authors:
  - name: Gavin L. Simpson
    orcid: 0000-0002-9084-8413
    affiliation: 1
affiliations:
 - name: Department of Animal and Veterinary Science, Aarhus University, Denmark
   index: 1
date: 26 June 2024
citation_author: Simpson
year: 2024
bibliography: paper.bib
output:
  rticles::joss_article:
    keep_md: true
csl: apa.csl
journal: JOSS
---

@@

Model diagnostic plots can be produced using `appraise()`, which by default produces four plots: i) a QQ plot of model residuals, with theoretical quantiles and reference bands generated following @Augustin2012-sc, ii) a plot of residuals (deviance residuals are the default) against linear predictor values, iii) a histogram of residuals, and iv) a plot of observed versus fitted values. Model diagnostic plots for the model, with simulated residuals-based reference bands on the QQ plot, are produced with

```{r m1-appraise, fig.show = "hide", dependson = "fit-chl-model", cache = FALSE, fig.height = 7, fig.width = 10, fig.pos = "t!"}
appraise(m1, method = "simulate")
```

which show significant heteroscedasticity and departure from the condtional distribution of the response given the model (Figure \ref{fig:m1-appraise}).
